\LARGE\bf Bayesian selection of best subsets in high-dimensional regression

### 2019-04-22 Manuscript/Jin/Project2.tex
set tau = log(p)^2

A possible drawback of our deterministic search is that the proposed algorithm can get trapped in a local optimum. To resolve this issue, we propose to add a stochastic global search algorithm to the proposed deterministic local search algorithm. This additional step will be used only to check whether or not the local search algorithm reaches the global optimum. To develop a fast global search algorithm, we employ a notion of the shotgun stochastic search \citep{hans2007shotgun}. However, while the shotgun stochastic search evaluates the candidate models using parallel computing, our simultaneous evaluation is based on analytical calculations. Our stochastic search algorithm can be implemented by generating a Markov chain via the following iterative sampling scheme:
\begin{itemize}
\item[1.] Given $\bg^*$, generate $\ueta^*$ with probability,
$$\pi(\ueta|\uy,\bg^*)=\frac{m(\uy|\ueta)}{\sum_{\ueta \in \mathcal{N}_+(\bg^* )  } m(\uy|\ueta) }\mathbb{I}\{\ueta \in \mathcal{N}_+(\bg^* ) \}.$$
\item[2.] Given $\ueta^*$, generate $\bg^*$ with probability
$$\pi(\bg|\uy,\ueta^*)=\frac{m(\uy|\bg)}{\sum_{\bg \in \mathcal{N}_-({\ueta}^* )  } m(\uy|\bg) }\mathbb{I}\{\bg \in \mathcal{N}_-({\ueta}^* )   \}.$$
\end{itemize}
Recall that  using $\eqref{eq:app:1}$ and  $\eqref{eq:app:2}$, calculating $\pi(\ueta|\uy,\bg^*)$ for all $\ueta \in \mathcal{N}_+(\bg^* )$ can be done simultaneously. Similarly, we can calculate $\pi(\bg|\uy,\ueta^*)$ for all $\bg \in \mathcal{N}_-({\ueta}^* ) $ at once.   
Using the notion of the Metropolis-Hasting algorithm, it is easy to check that the states of generated Markov chain by our sampling method is equivalent to the support of $\pi(\bg|\uy)$\textcolor{red}{(Revise it later)}. Let $\bg^{(t)}$ be the sample generated at iteration $t$. If the outcome from Algorithm \ref{alg:1}, $\hat{\bg}$, is not the global maximum, then we must observe at least one case that $\pi(\hat{\bg}|\uy)< \pi(\bg^{(t)}|\uy)$ (or equivalently $m(\uy|\hat{\bg})<m(\uy|\bg^{(t)})$) with probability one as $t\to \infty$. Therefore, if this case does not occur for sufficiently large $t$, we can treat $\hat{\bg}$ as the global maximum.

\begin{algorithm}[ht]
\caption{Hybrid best subset search with a fixed $k$}\label{alg:2}
\begin{itemize}
\item[1.] Initialize $\tilde{\bg}$ and $\tilde{\tau}$.
\item[2.] \textbf{Repeat}
\begin{itemize}
\item[] Update $\tilde{\ueta}\leftarrow \arg\max_{\ueta  \in \mathcal{N}_+(\tilde{\bg} ) } m(\uy|\ueta,\tilde{\tau});$ 
\item[] Update $\tilde{\bg}\leftarrow  \arg\max_{\bg  \in \mathcal{N}_-(\tilde{\ueta} ) } m(\uy|\bg,\tilde{\tau});$
\item[] Update $\tilde{\tau} \leftarrow \arg\min_{\tau \in (0,\infty)}Q(\tau|\tilde{\bg})$;
\end{itemize}
\item[] \textbf{until} convergence.
\item[3.] Set $\bg^{(0)}=\tilde{\bg}$.
\item[4.] \textbf{Repeat} for $t=1,\ldots,T$:
\begin{itemize}
\item[] Generate $\ueta^{(t)}\sim \pi(\ueta|\uy,\bg^{(t-1)})$;
\item[] Generate $\bg^{(t)}\sim \pi(\bg|\uy,\ueta^{(t)})$;
\item[] \textbf{If} $m(\uy|\tilde{\bg})<m(\uy|\bg^{(t)})$, \textbf{then} set $\tilde{\bg}=\bg^{(t)}$, break the loop, and go to 2.
\end{itemize}
\item[5.] Return $\hat{\bg}=\tilde{\bg}$.
\end{itemize}
\end{algorithm}

@@

\begin{table}[H]
  \centering
  \caption{Simulation study results based on $2,000$ replications; FDR (false discovery rate), TRUE\% (percentage of the true model detected), SIZE (selected model size), HAM (Hamming distance), and s.e. (standard error).}\label{T:sim1}
  \begin{tabular}{cc|c|c|c|c}
    \hline
    Scenario&Method&FDR  (s.e.) & TRUE\% (s.e.) & SIZE (s.e.) & HAM (s.e.)  \\
    \hline
i&Proposed & 0.016 (0.001) & 92.600 (0.585)& 4.087 (0.007) & 0.087 (0.007) \\ 
    &SCAD & 0.034 (0.002) &  85.200 (0.794) & 4.188 (0.011) & 0.188 (0.011)\\ 
   &MCP & 0.035 (0.002) & 84.750 (0.804) & 4.191 (0.011) & 0.191 (0.011)\\ 
  &ENET & 0.016 (0.001) & 92.700 (0.582) & 4.087 (0.007) & 0.087 (0.007)\\ 
  &LASSO & 0.020 (0.002) & 91.350 (0.629) & 4.109 (0.009) & 0.109 (0.009)\\
    \hline
    ii&Proposed & 0.030 (0.002) & 86.250 (0.770) & 4.034 (0.007) & 0.236 (0.014)   \\
&SCAD & 0.059 (0.003) & 74.150 (0.979) & 4.107 (0.015) & 0.480 (0.022) \\ 
  &MCP & 0.137 (0.004) & 55.400 (1.112) & 4.264 (0.020) & 1.098 (0.034) \\ 
  &ENET & 0.501 (0.004) & 0.300 (0.122) & 7.716 (0.072) & 5.018 (0.052) \\ 
  &LASSO & 0.276 (0.004) & 15.550 (0.811) & 5.308 (0.033) & 2.038 (0.034) \\ 
    \hline
    iii&Proposed&0.014 (0.001) & 93.450 (0.553) & 4.073 (0.007) & 0.073 (0.007)\\
&SCAD & 0.027 (0.002) & 87.900 (0.729) & 4.145 (0.010) & 0.145 (0.010) \\ 
  &MCP & 0.031 (0.002) & 86.550 (0.763) & 4.172 (0.013) & 0.172 (0.013) \\ 
  &ENET & 0.035 (0.002) & 84.850 (0.802) & 4.181 (0.013) & 0.206 (0.012) \\ 
  &LASSO & 0.014 (0.001) & 93.850 (0.537) & 4.073 (0.007) & 0.073 (0.007) \\ 
    \hline
    iv&Proposed&0.032 (0.002) & 86.700 (0.759) & 4.051 (0.007) & 0.236 (0.015) \\
&SCAD & 0.068 (0.003) & 74.250 (0.978) & 4.196 (0.014) & 0.493 (0.023) \\ 
  &MCP & 0.152 (0.004) & 53.750 (1.115) & 4.226 (0.017) & 1.202 (0.035) \\ 
  &ENET & 0.417 (0.005) & 0.150 (0.087) & 6.228 (0.068) & 4.089 (0.043) \\ 
  &LASSO & 0.265 (0.004) & 19.500 (0.886) & 5.139 (0.029) & 1.909 (0.035) \\ 
    \hline
  \end{tabular}
\end{table}

\section{Real data application}
 In this section, we apply our proposed methods to breast cancer tissue samples from The Cancer Genome Atlas (TCGA) Research Network: \url{http://cancergenome.nih.gov}. The dataset "Breast Invasive Carcinoma" (BRCA) we use in this paper is extracted from the R package "curatedTCGAData" and is publicly available and described at \url{http://bioconductor.org/packages/release/data/experiment/html/curatedTCGAData.html}, where includes many other Bioconductor MultiAssayExperiment class objects. The data set contains $17814$ gene expression measurements originally. After removing missing data, it is left $526$ patients with $17,323$ gene expression measurements (recorded on the log scale).  BRCA1 is a tumor suppressor gene and its mutations predispose women to breast cancer \citep{findlay2018accurate}. Our goal here is to identify the best fitting model for estimating an association between BRCA1 (response variable) and the other genes (independent variables). We standardize all the variables.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & NSP & MSE\_Pred & MSE\_fit & BIC & EBIC & Detection Rate\\ 
  \hline
  Local & 31.00 & 0.48 & 0.40 & 738.91 & 1048.20 & 2/11\\ 
  Grobal & 50.00 & 0.39 & 0.30 & 548.03 & 1015.81 & 1/11\\ 
  SCAD & 4.00 & 0.69 & 0.67 & 1104.69 & 1154.11 & 1/11\\ 
  MCP & 4.00 & 0.69 & 0.67 & 1104.69 & 1154.11 & 1/11\\ 
  ENET & 5.00 & 0.69 & 0.67 & 1110.65 & 1171.30 & 1/11\\ 
  LASSO & 4.00 & 0.69 & 0.67 & 1104.69 & 1154.11 & 1/11\\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}
  \includegraphics[width=\linewidth]{Heatmap.pdf}
  \caption{Heap map}
  \label{fig:heatmap}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{Log_marginal.pdf}
  \caption{Logmarginal}
  \label{fig:Logmarginal}
\end{figure}

>>>>

A possible drawback of our deterministic search is that the proposed algorithm can get trapped in a local optimum. To resolve this issue, we propose to add a stochastic global search algorithm to the proposed deterministic local search algorithm. This additional step will be used only to check whether or not the local search algorithm reaches the global optimum. To develop a fast global search algorithm, we employ a notion of the shotgun stochastic search \citep{hans2007shotgun}. However, while the shotgun stochastic search evaluates the candidate models using parallel computing, our simultaneous evaluation is based on analytical calculations. Our stochastic search algorithm can be implemented by generating a Markov chain via the following iterative sampling scheme:
 \item[1.] Given $\bg^*$, generate $\ueta^*$ with probability,
       $$\pi(\ueta|\uy,\bg^*)=\frac{m(\uy|\ueta)}{\sum_{\ueta \in \mathcal{N}_+(\bg^* )  } m(\uy|\ueta) }\mathbb{I}\{\ueta \in \mathcal{N}_+(\bg^* ) \}.$$
 \item[2.] Given $\ueta^*$, generate $\bg^*$ with probability
       $$\pi(\bg|\uy,\ueta^*)=\frac{m(\uy|\bg)}{\sum_{\bg \in \mathcal{N}_-({\ueta}^* )  } m(\uy|\bg) }\mathbb{I}\{\bg \in \mathcal{N}_-({\ueta}^* )   \}.$$
Recall that  using $\eqref{eq:app:1}$ and  $\eqref{eq:app:2}$, calculating $\pi(\ueta|\uy,\bg^*)$ for all $\ueta \in \mathcal{N}_+(\bg^* )$ can be done simultaneously. Similarly, we can calculate $\pi(\bg|\uy,\ueta^*)$ for all $\bg \in \mathcal{N}_-({\ueta}^* ) $ at once.
Using the notion of the Metropolis-Hasting algorithm, it is easy to check that the states of generated Markov chain by our sampling method is equivalent to the support of $\pi(\bg|\uy)$\textcolor{red}{(Revise it later)}. Let $\bg^{(t)}$ be the sample generated at iteration $t$. If the outcome from Algorithm \ref{alg:1}, $\hat{\bg}$, is not the global maximum, then we must observe at least one case that $\pi(\hat{\bg}|\uy)< \pi(\bg^{(t)}|\uy)$ (or equivalently $m(\uy|\hat{\bg})<m(\uy|\bg^{(t)})$) with probability one as $t\to \infty$. Therefore, if this case does not occur for sufficiently large $t$, we can treat $\hat{\bg}$ as the global maximum.
\textcolor{blue}{However, this global maximum is unefficient because it require large enough $t$. To improve the efficiency of global search, We applied the idea from Simulated Annealing to Algorithm \ref{alg:2}:
 \item[1.] Given $\bg^*$, generate $\ueta^*$ with probability,
       $$\pi(\ueta|\uy,\bg^*)=\frac{\tilde{m}(\uy|\ueta)}{\sum_{\ueta \in \mathcal{N}_+(\bg^* )  } \tilde{m}(\uy|\ueta) }\mathbb{I}\{\ueta \in \mathcal{N}_+(\bg^* ) \},$$ where $\tilde{m}(\uy|\ueta)=\exp\{\frac{1}{\alpha}\log(m(\uy|\ueta)))\}.$ \\
 \item[2.] Given $\ueta^*$, generate $\bg^*$ with probability
       $$\pi(\bg|\uy,\ueta^*)=\frac{\tilde{m}(\uy|\bg)}{\sum_{\bg \in \mathcal{N}_-({\ueta}^* )  } \tilde{m}(\uy|\bg) }\mathbb{I}\{\bg \in \mathcal{N}_-({\ueta}^* )  \},$$ where $\tilde{m}(\uy|\bg)=\exp\{\frac{1}{\alpha}\log(m(\uy|\bg))\}.$
}
\begin{algorithm}[ht]
 \caption{Hybrid best subset search with a fixed $k$}\label{alg:2}
  \item[1.] Initialize $\tilde{\bg}$ and $\tilde{\tau}$.
  \item[2.] \textbf{Repeat}
         \item[] Update $\tilde{\ueta}\leftarrow \arg\max_{\ueta  \in \mathcal{N}_+(\tilde{\bg} ) } m(\uy|\ueta,\tilde{\tau});$
         \item[] Update $\tilde{\bg}\leftarrow  \arg\max_{\bg  \in \mathcal{N}_-(\tilde{\ueta} ) } m(\uy|\bg,\tilde{\tau});$
  \item[] \textbf{until} convergence.
  \item[3.] Set $\bg^{(0)}=\tilde{\bg}$.
  \item[4.] \textbf{Repeat} for $t=1,\ldots,T$:
         \item[] Generate $\ueta^{(t)}\sim \pi(\ueta|\uy,\bg^{(t-1)})$;
         \item[] Generate $\bg^{(t)}\sim \pi(\bg|\uy,\ueta^{(t)})$;
         \item[] \textbf{If} $m(\uy|\tilde{\bg})<m(\uy|\bg^{(t)})$, \textbf{then} set $\tilde{\bg}=\bg^{(t)}$, break the loop, and go to 2.
  \item[5.] Return $\hat{\bg}=\tilde{\bg}$.
\end{algorithm}

@@

\begin{table}[H]
 \centering
 \caption{Simulation study results based on $2,000$ replications; FDR (false discovery rate), TRUE\% (percentage of the true model detected), SIZE (selected model size), HAM (Hamming distance), and s.e. (standard error).}\label{T:sim1}
 \begin{tabular}{cc|c|c|c|c}
  \hline
  Scenario & Method                         & FDR  (s.e.)    & TRUE\% (s.e.)   & SIZE (s.e.)     & HAM (s.e.)    \\
  \hline
  i        & Proposed                       & 0.016 (0.001)  & 92.600 (0.585)  & 4.087 (0.007)   & 0.087 (0.007) \\
           & \textcolor{red}{local(new)}    & 0.01(0.001)    & 96.90 (0.388)   & 4.03  (0.004)   & 0.03 (0.004)  \\
           & \textcolor{red}{Proposed(new)} & 0.006 (0.001)  & 96.900 (0.388)  & 4.032(0.004)    & 0.032 (0.004) \\
           & SCAD                           & 0.034 (0.002)  & 85.200 (0.794)  & 4.188 (0.011)   & 0.188 (0.011) \\
           & MCP                            & 0.035 (0.002)  & 84.750 (0.804)  & 4.191 (0.011)   & 0.191 (0.011) \\
           & ENET                           & 0.016 (0.001)  & 92.700 (0.582)  & 4.087 (0.007)   & 0.087 (0.007) \\
           & LASSO                          & 0.020 (0.002)  & 91.350 (0.629)  & 4.109 (0.009)   & 0.109 (0.009) \\
  \hline
  ii       & Proposed                       & 0.030 (0.002)  & 86.250 (0.770)  & 4.034 (0.007)   & 0.236 (0.014) \\
           & \textcolor{red}{local(new)}    & 0.03  (0.002)  & 86.15  (0.773)  & 3.99 (0.007)    & 0.25  (0.015) \\
           & \textcolor{red}{Proposed(new)} & 0.023  (0.002) & 88.750  (0.707) & 3.985  (0.006)  & 0.203 (0.014) \\
           & SCAD                           & 0.059 (0.003)  & 74.150 (0.979)  & 4.107 (0.015)   & 0.480 (0.022) \\
           & MCP                            & 0.137 (0.004)  & 55.400 (1.112)  & 4.264 (0.020)   & 1.098 (0.034) \\
           & ENET                           & 0.501 (0.004)  & 0.300 (0.122)   & 7.716 (0.072)   & 5.018 (0.052) \\
           & LASSO                          & 0.276 (0.004)  & 15.550 (0.811)  & 5.308 (0.033)   & 2.038 (0.034) \\
  \hline
  iii      & Proposed                       & 0.014 (0.001)  & 93.450 (0.553)  & 4.073 (0.007)   & 0.073 (0.007) \\
           & \textcolor{red}{local(new)}    & 0.00  (0.001)  & 98.10  (0.305)  & 4.02  (0.003)   & 0.02(0.003)   \\
           & \textcolor{red}{Proposed(new)} & 0.004(0.001)   & 98.100 (0.305)  & 4.020(0.003)    & 0.020 (0.003) \\
           & SCAD                           & 0.027 (0.002)  & 87.900 (0.729)  & 4.145 (0.010)   & 0.145 (0.010) \\
           & MCP                            & 0.031 (0.002)  & 86.550 (0.763)  & 4.172 (0.013)   & 0.172 (0.013) \\
           & ENET                           & 0.035 (0.002)  & 84.850 (0.802)  & 4.181 (0.013)   & 0.206 (0.012) \\
           & LASSO                          & 0.014 (0.001)  & 93.850 (0.537)  & 4.073 (0.007)   & 0.073 (0.007) \\
  \hline
  iv       & Proposed                       & 0.032 (0.002)  & 86.700 (0.759)  & 4.051 (0.007)   & 0.236 (0.015) \\
           & \textcolor{red}{local(new)}    & 0.04  (0.002)  & 85.30  (0.792)  & 4.02  (0.006)   & 0.28  (0.016) \\
           & \textcolor{red}{Proposed(new)} & 0.023(0.002)   & 89.850 (0.675)  & 4.005  (0.005 ) & 0.190 (0.013) \\
           & SCAD                           & 0.068 (0.003)  & 74.250 (0.978)  & 4.196 (0.014)   & 0.493 (0.023) \\
           & MCP                            & 0.152 (0.004)  & 53.750 (1.115)  & 4.226 (0.017)   & 1.202 (0.035) \\
           & ENET                           & 0.417 (0.005)  & 0.150 (0.087)   & 6.228 (0.068)   & 4.089 (0.043) \\
           & LASSO                          & 0.265 (0.004)  & 19.500 (0.886)  & 5.139 (0.029)   & 1.909 (0.035) \\
  \hline
 \end{tabular}
\end{table}

\section{Real data example using mRNA gene expression data}
\subsection{Data and data precessing}
"Breast Invasive Carcinoma" (BRCA) is mRNA gene expression dataset from the R package "curatedTCGAData" in \citet{Ramos2019}. This package provides us with publicly available data from The Cancer Genome Atlas (TCGA) Bioconductor MultiAssayExperiment class objects. BRCA dataset comprises $526$ patients with primary solid tumor and $17814$ gene expression measurements, including a gene called BRCA1. BRCA1 is a well-known tumor suppressor gene and its mutations predispose women to breast cancer \citep{findlay2018accurate}. Our goal here is to identify the best fitting model for estimating an association between BRCA1 (response variable) and the other genes (independent variables).

After removing some missing data, we have $n=526$ samples with $17,323$ genes. Then it is further screened and normalized and we selected $p=5000$ genes that are most marginally correlated with the response BRCA1 gene expression measurement.
\subsection{Results}
In our proposed method, the upper bound of subset size $K=\lceil n^{2/3} \rceil =66$, $\tau=(\log p)^2=72.54$. To evaluate the fit of model, we computed the average of prediction MSE, BIC and EBIC under the our proposed method compared with SCAD, MCP, elastic net and LASSO. We used cross validation method to randomly split data into $70\%$ training set and remaining part as testing set. For selected genes by each method, we fitted the linear model and compute the MSE, BIC and EBIC. Repeated this procedure $500$ times and compute the average.

Table \ref{T:realdata} shows our proposed method selects $8$ genes (NSP: number of selected predictors) while others $4$ or $5$. In addition, it indicates that our proposed method outperforms other models since it has smallest values of prediction MSE, BIC and EBIC. Fig \ref{Fig:heatmap} gives us a heat map of the specific gene's names selected and the magnitude of coefficients estimated by each methods under the linear regression model with the whole data. "X" represents non-significance of coefficient.

Fig \ref{Fig:logmarginal} is $\log m(\bg|\uy)$ with $k$ predictors selected by proposed method. It increases to be maximized at when $k=8$ and then decreases all the way down.
\begin{table}[ht]
 \centering
 \caption{Predicion MSE, BIC and EBIC of Proposed method and other methods}\label{T:realdata}
 \begin{tabular}{rrrrr}
  \hline
           & NSP  & MSE\_Pred & BIC     & EBIC    \\
  \hline
  Proposed & 8.00 & 0.60      & 984.45  & 1099.50 \\
  SCAD     & 4.00 & 0.68      & 1104.69 & 1166.47 \\
  MCP      & 4.00 & 0.68      & 1104.69 & 1166.47 \\
  ENET     & 5.00 & 0.68      & 1110.65 & 1186.25 \\
  LASSO    & 4.00 & 0.68      & 1104.69 & 1166.47 \\
  \hline
 \end{tabular}
\end{table}

\begin{figure}
 \centering
 \includegraphics{Heatmap.pdf}
 \caption{Heap map}\label{Fig:heatmap}
 \label{fig:heatmap}
\end{figure}

\section{Comments}
We have presented a novel approach for best subsets selection, a hybrid algorithm for exploring model spaces via local search and stochastic global search to consider and evaluate thousands of possible predictors simultaneously. The ability of identification of model spaces rapidly and accurately will be more practical than some traditional methods, such as MCMC stochastic search, which are inefficient due to high dimensions and high noise data. Furthermore, this approach can be applied into multivariate response regression model and binary regression model with logit link and probit link since marginal likelihood of them can be assessed or approximated quickly.

In future, we plan to introduce distributed computing (or parallel computing) to further improve the speed, e.g. evaluating best subset for each $k$ by parallel computing. Besides, we plan to apply this algorithm into multivariate linear regression model, probit model and then multivariate mixed response regression model so that this model can be used in clinical data with both continuous and binary responses data.
